---
title: "Spotify data exploration and modeling"
subtitle: "Or, how to make the perfect song"
author: "The IDDS 2201 team"
date: "2022-04-06"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r package load in, message=F,warning=F, echo=F}
library(tidyverse)
library(tidymodels)
```



Today we will explore our Spotify playlist data! 

We will use this as an opportunity to practice data: 

- import
- exploration
- transformation and tidying
- visualization,
- and functional and iterative programming.

We will also take this opportunity to learn: 

- modeling and
- model interpretation.

Our goal here is to find **which variables explain the most variation in track popularity.**

For reference, you can always look at our [class lecture notes](https://vcannataro.github.io/IDDS_2201_public/Lectures/), and the lecture slide deck on [modeling](https://vcannataro.github.io/IDDS_2201_public/Lectures/L12_modeling_p1/L12_modeling_part_1.html?panelset=plot&panelset1=plot2&panelset2=plot3&panelset3=plot4&panelset4=plot5&panelset5=plot6&panelset6=plot7&panelset7=plot8#34) is especially useful! 

Let's start with importing our class playlist and seeing what the data looks like. Head over to the `data/` subdirectory and find the `IDDS_class_playlist`. Look at the `README.md` file to see definitions of some variables in our dataset. 


# Data import and check

```{r import class playlist}
# What type of data file is the IDDS class playlist? 
# 
```


# Data explore 

# Data visualize and model




