---
title: "Spotify data exploration and modeling"
subtitle: "Or, how to make the perfect song"
author: "The IDDS 2201 team"
date: "2022-04-06"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r package load in, message=F,warning=F, echo=F}
library(tidyverse)
library(tidymodels)
```



Today we will explore our Spotify playlist data! 

We will use this as an opportunity to practice data: 

- import
- exploration
- transformation and tidying
- visualization,
- and functional and iterative programming.

We will also take this opportunity to learn: 

- modeling and
- model interpretation.

Our goal here is to find **which variables explain the most variation in track popularity.**

For reference, you can always look at our [class lecture notes](https://vcannataro.github.io/IDDS_2201_public/Lectures/), and the lecture slide deck on [modeling](https://vcannataro.github.io/IDDS_2201_public/Lectures/L12_modeling_p1/L12_modeling_part_1.html?panelset=plot&panelset1=plot2&panelset2=plot3&panelset3=plot4&panelset4=plot5&panelset5=plot6&panelset6=plot7&panelset7=plot8#34) is especially useful! 

Let's start with importing our class playlist and seeing what the data looks like. Head over to the `data/` subdirectory and find the `IDDS_class_playlist`. Look at the `README.md` file to see definitions of some variables in our dataset. 


# Data import and check

```{r import class playlist}
# What type of data file is the IDDS class playlist? 

IDDS_class_playlist <- read_rds(file = "data/IDDS_class_playlist.rds")

```


# Data explore 

```{r}
glimpse(IDDS_class_playlist)
```



```{r}
IDDS_class_playlist %>%
  ggplot(aes(x=track.popularity)) + 
  geom_histogram()
```





# Data visualize and model


```{r}
IDDS_class_playlist %>% 
  ggplot(aes(y=track.popularity, x=danceability)) + 
  geom_point() + 
  geom_smooth(method = "lm")
```

```{r}
dance_regression <- linear_reg() %>%
  set_engine("lm") %>%
  fit(track.popularity ~ danceability, data = IDDS_class_playlist)

parsnip::tidy(dance_regression)

glance(dance_regression)

```

# building a function


```{r}

variable_of_interest <- "danceability"

lm_stats <- function(variable_of_interest){
  
  response_variable <- IDDS_class_playlist %>%
    pull(track.popularity)
  
  explanatory_variable <- IDDS_class_playlist %>%
    pull(variable_of_interest) 
  
  new_df <- data.frame(response_variable, explanatory_variable)
  
  new_fit <- linear_reg() %>%
    set_engine("lm") %>% 
    fit(response_variable ~ explanatory_variable, data = new_df)
  
  slope <- parsnip::tidy(new_fit) %>%
    filter(term == "explanatory_variable") %>%
    pull(estimate)
  
  r_squared <- glance(new_fit) %>%
    pull(r.squared)
  
  p_value <- glance(new_fit) %>%
    pull(p.value)
  
  return(
    list(
      slope = slope, 
      r_squared = r_squared,
      p_value = p_value
      )
    
    )
}


lm_stats(variable_of_interest = "danceability")
lm_stats(variable_of_interest = "valence")
lm_stats(variable_of_interest = "key")

```


```{r}
my_variables <- IDDS_class_playlist %>%
  select_if(is.numeric) %>%
  select(!track.popularity) %>% 
  names()

IDDS_stats <- data.frame(my_variables, slope=NA, r_squared = NA, p_value = NA)


row_ind <- 1


for(row_ind in 1:nrow(IDDS_stats)  ){
  
  stats_results <- lm_stats(variable_of_interest = IDDS_stats[row_ind,"my_variables"])
  
  IDDS_stats[row_ind, "slope"] <- stats_results$slope
  
  IDDS_stats[row_ind, "r_squared"] <- stats_results$r_squared
  
  IDDS_stats[row_ind, "p_value"] <- stats_results$p_value
  
  message("I finished this loop")
  
}

IDDS_stats
```





```{r}
spotify_big_data <- read_csv("data/spotify_big_data.csv") 
glimpse(spotify_big_data)
```






```{r}

lm_stats_spotify <- function(variable_of_interest){
  
  response_variable <- spotify_big_data %>%
    pull(track.popularity)
  
  explanatory_variable <- spotify_big_data %>%
    pull(variable_of_interest) 
  
  new_df <- data.frame(response_variable, explanatory_variable)
  
  new_fit <- linear_reg() %>%
    set_engine("lm") %>% 
    fit(response_variable ~ explanatory_variable, data = new_df)
  
  slope <- parsnip::tidy(new_fit) %>%
    filter(term == "explanatory_variable") %>%
    pull(estimate)
  
  r_squared <- glance(new_fit) %>%
    pull(r.squared)
  
  p_value <- glance(new_fit) %>%
    pull(p.value)
  
  return(
    list(
      slope = slope, 
      r_squared = r_squared,
      p_value = p_value
      )
    
    )
}



my_variables_big <- spotify_big_data %>%
  select_if(is.numeric) %>%
  select(!track.popularity) %>% 
  names()

spotify_stats <- data.frame(my_variables_big, slope=NA, r_squared = NA, p_value = NA)


row_ind <- 1


for(row_ind in 1:nrow(spotify_stats)  ){
  
  stats_results <- lm_stats_spotify(variable_of_interest = spotify_stats[row_ind,"my_variables_big"])
  
  spotify_stats[row_ind, "slope"] <- stats_results$slope
  
  spotify_stats[row_ind, "r_squared"] <- stats_results$r_squared
  
  spotify_stats[row_ind, "p_value"] <- stats_results$p_value
  
  message("I finished this loop")
  
}

spotify_stats


```





